A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-08 paper/paper.Rmd
revised target data section

The NLSY79 data used in @SingerJudithD2003Alda contains the longitudinal measurements of the male high school dropouts who first participated in the study at age 14-17 years from 1979 through to 1994. This dataset comprises of several variables as follows:

1. ID: the respondents' ID. 
2. EXPER: temporal scale, i.e., the length of time (years) in the workforce, starting on the respondents' first day at work.
3. LNW: natural logarithm of wages, adjusted with 1990's inflation rate. 
4. BLACK: binary variables, the respondents are coded to 1 because they were Black and 0 otherwise. 
5. HGC: the highest grade completed.
6. UERATE: unemployment rate of the year of survey. When missing, the variable is set to be 7.875 (the average rate).

From the database, we will refresh this data with a more up to date period, which is from survey year 1979 to survey year 2018 (the most recent year published). We will  also modify the variables (for example, we create a race variable, instead of the binary variable "Black") and add some variables so that more examples of data exploration can be displayed in teaching. The variables added are year of survey, age of individual in 1979, year when the highest grade completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual starting to work, and the number of years the individual worked.

In addition to the dropouts data, we will also make the up to high school graduates' wages available, including the females, and the demographic data of the NLSY79 cohort available in separate datasets. Thus, we will create three datasets as follows:

1. The demographic data of the NLSY79 cohort.
2. The wages data of the up to high school graduates.
3. The wages data of the high school dropouts, which is the refreshed version of @SingerJudithD2003Alda's data.

>>>>

The NLSY79 data used in @SingerJudithD2003Alda contains the longitudinal records of male high school dropouts who first participated in the study at age 14-17 years from 1979 through to 1994. This dataset contains several variables as follows:

1. ID: the respondents' ID. 
2. EXPER: temporal scale, i.e., the length of time (years) in the workforce, starting on the respondents' first day at work.
3. LNW: natural logarithm of wages, adjusted with 1990's inflation rate. 
4. BLACK: binary variable, 1 indicates black and 0 otherwise. 
5. HISPANIC: binary variable, 1 indicates hispanic and 0 otherwise. 
6. HGC: the highest grade completed.
7. UERATE: unemployment rate of the year of survey. When missing, the variable is set to be 7.875 (the average rate).

From the database, we will refresh this data by re-creating the full data with records from survey years 1979 through to 2018 (the most recent year published). We also plan to make some modifications of the variables. For example, we create a categorical race variable instead of the binary variable "black" and "hispanic".  We also include additional variables so that more interesting data exploration can be included in teaching examples: year of survey, age of individual in 1979, year when the highest grade was completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual started to work, and the number of years the individual worked.

We will also make all of the high school graduates data available, including females, in addition to high school dropouts, and separate the demographic data of the NLSY79 cohort. Thus, the plan is to create three datasets as follows:

1. The demographic data of the NLSY79 cohort.
2. The wages data of the up to high school graduates.
3. The wages data of the high school dropouts, which is closest to a refreshed version of @SingerJudithD2003Alda's data.

### 2022-02-08 paper/paper.tex
revised target data section

The NLSY79 data used in Singer and Willett (2003) contains the longitudinal measurements of the male high school dropouts who first participated in the study at age 14-17 years from 1979 through to 1994. This dataset comprises of several variables as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  ID: the respondents' ID.
\item
  EXPER: temporal scale, i.e., the length of time (years) in the workforce, starting on the respondents' first day at work.
\item
  LNW: natural logarithm of wages, adjusted with 1990's inflation rate.
\item
  BLACK: binary variables, the respondents are coded to 1 because they were Black and 0 otherwise.
\item
  HGC: the highest grade completed.
\item
  UERATE: unemployment rate of the year of survey. When missing, the variable is set to be 7.875 (the average rate).
\end{enumerate}

From the database, we will refresh this data with a more up to date period, which is from survey year 1979 to survey year 2018 (the most recent year published). We will also modify the variables (for example, we create a race variable, instead of the binary variable ``Black'') and add some variables so that more examples of data exploration can be displayed in teaching. The variables added are year of survey, age of individual in 1979, year when the highest grade completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual starting to work, and the number of years the individual worked.

In addition to the dropouts data, we will also make the up to high school graduates' wages available, including the females, and the demographic data of the NLSY79 cohort available in separate datasets. Thus, we will create three datasets as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The demographic data of the NLSY79 cohort.
\item
  The wages data of the up to high school graduates.
\item
  The wages data of the high school dropouts, which is the refreshed version of Singer and Willett (2003)'s data.
\end{enumerate}

>>>>

The NLSY79 data used in Singer and Willett (2003) contains the longitudinal records of male high school dropouts who first participated in the study at age 14-17 years from 1979 through to 1994. This dataset contains several variables as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  ID: the respondents' ID.
\item
  EXPER: temporal scale, i.e., the length of time (years) in the workforce, starting on the respondents' first day at work.
\item
  LNW: natural logarithm of wages, adjusted with 1990's inflation rate.
\item
  BLACK: binary variable, 1 indicates black and 0 otherwise.
\item
  HISPANIC: binary variable, 1 indicates hispanic and 0 otherwise.
\item
  HGC: the highest grade completed.
\item
  UERATE: unemployment rate of the year of survey. When missing, the variable is set to be 7.875 (the average rate).
\end{enumerate}

From the database, we will refresh this data by re-creating the full data with records from survey years 1979 through to 2018 (the most recent year published). We also plan to make some modifications of the variables. For example, we create a categorical race variable instead of the binary variable ``black'' and ``hispanic''. We also include additional variables so that more interesting data exploration can be included in teaching examples: year of survey, age of individual in 1979, year when the highest grade was completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual started to work, and the number of years the individual worked.

We will also make all of the high school graduates data available, including females, in addition to high school dropouts, and separate the demographic data of the NLSY79 cohort. Thus, the plan is to create three datasets as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The demographic data of the NLSY79 cohort.
\item
  The wages data of the up to high school graduates.
\item
  The wages data of the high school dropouts, which is closest to a refreshed version of Singer and Willett (2003)'s data.
\end{enumerate}
